Feasibility of Harvesting Power To Run A Domestic Water Meter Using Streaming Cell Technology

### 2014-08-06 paper.tex
Added content to water consumption estimation

    A double layer can be broken into two sub-layers; the Stern and diffuse layers.
    The Stern layer is comprised of immobilised ions situated very close the charged solid surface.~\cite{Salieb-Beugelaar2009}
    The diffuse layer is formed by the ions attracted to the surface, but not so strongly as to be immobile.

    A streaming cell makes use of ions at the interface by forcing them through one or more narrow channels.
    The channels can be formed from porous materials or created individually by various fabrication methods, e.g. chemical etching.
    By immersing a channel in an electrolyte solution, double layers line the interior to the channel.
    The thickness of the double layer is represented by the Debye length ($l_{D}$) and is controlled the ionic concentration of the liquid.~\cite{Israelachvili2011}
    When the dimensions of a channel become small enough the double layers begin to overlap and co-ions are repelled from the channel.

    Counter-ions within a streaming cell are collected by applying a pressure gradient across the channel/cell.
    Pressure forces the liquid through the channel (from which co-ions have been removed) creating a streaming current.
    It is the streaming current that raises the voltage (streaming potential) across the cell as the ions are pumped through it.

    Streaming currents have been heavily investigated as a means of generating electrical energy from pressure gradients.~\cite{Chang2009,Daiguji2006,Daiguji2004b,Davidson2008a,Davidson2008,CherngHon2012,Jiao2014,Lu2006,Olthuis2005,Osterle1964,Pennathur2007,Ren2008a,VanderHeyden2006,Heyden2007,Xie2008,Yang2003}
    % It is expected that a channel height equal to twice the Debye length leads to maximum energy conversion efficiency.~\cite{Daiguji2004b}
    Theoretical predictions of the efficiency of standard micro/nano-fluidic channels are 2\% for pure water and 7\% for sodium chloride.~\cite{VanderHeyden2006}
    Experimental results show conversion efficiencies in the range of:
    \begin{itemize}
        \item 0.01\% by forcing water through porous glass with pore sizes from 10\thinspace--\SI{16}{\micro\metre}.~\cite{Yang2003}
        \item 0.8\% by forcing pure water through a ceramic rod populated with \SI{6}{\micro\metre} pores.~\cite{Yang2004}
        \item 3\% by forcing sodium chloride solution through a \SI{75}{\nano\metre} by \SI{50}{\micro\metre} silica channel.~\cite{Heyden2007}
        \item 0.77\% by forcing sodium chloride through a \SI{200}{\nano\metre} pore in an alumina membrane.~\cite{Lu2006}
        \item 5\% by forcing sodium chloride solution through a \SI{0.5}{\nano\metre} cylindrical pore in polyethylene terephalate foil.~\cite{Xie2008}
    \end{itemize}

@@

    The remainder of this work is presented as follows.
    In section \ref{sect:streamingCell} we build and measure a simple streaming potential cell.
    We confirm the relationship between pressure and applied pressure and investigate the effect of channel height.
    In section \ref{sect:waterConsumption} we estimate a typical domestic water usage profile.
    This consumption profile will be used as the basis for estimating the amount of energy available to the harvester.
    Section \ref{sect:powerRequirements} determines the amount of energy an electronic meter would need to be useful.
    Following on this, section \ref{sect:harvesterSize} presents the required dimensions a usable harvester would need.
    Finally we discuss the feasibility of using in the conclusion.

@@

    Each channel is epoxied between two acrylic reservoirs and a base plate.
    This provides a rigid structure under pressure and provides connection of wires and hoses.
    Copper wires and silicon hoses are inserted into each reservoir to facilitate streaming potential measurement and f:w

    A photo of one of the cells is shown in Fig. \ref{fig:cell}.

    \begin{tabular}{|c|c|c|} \hline
    Microscope slides & Sail Brand & JIA 7101WT - 26 x 76mm\tabularnewline \hline
    Shims & Garlock & Colorplast - 50$\,\mu$m, 80$\,\mu$m, 120$\,\mu$m and 250$\,\mu$m\tabularnewline \hline
    Epoxy & Selleys & Araldite - Ultra Clear Resin\tabularnewline \hline
    Pressure Sensor & Honeywell & 24PC15SMT - 0 -- $\pm$15 PSI\tabularnewline \hline
    \end{tabular}

    \section{Estimation of Typical Household Water Consumption}:
    \label{sect:waterConsumption}
    \section{Estimation of Power Requirements}
    \label{sect:powerRequirements}
    \section{Estimation of Streaming Cell Harvester}
    \label{sect:harvesterSize}
    \section{Conclusion}
    \bibliographystyle{plain}
    \bibliography{library}

>>>>

    A double layer can be broken into two sub-layers; the Stern and diffuse layers.
    The Stern layer is comprised of immobilised ions situated very close the charged solid surface.~\cite{Salieb-Beugelaar2009}
    The diffuse layer is formed by the ions attracted to the surface, but not so strongly as to be immobile.
    A streaming cell makes use of ions at the interface by forcing them through one or more narrow channels.

    The channels can be formed en-mass from using porous materials or created individually using a range fabrication methods, e.g. chemical etching.
    When using glass as the channel material, the surface spontaneously develops a negative surface charge when it comes in contact with water.
    This negative charge is caused by deprotonation at the wall from SiOH~$\leftrightarrows$~SiO$^{-}+$~H$^{+}$.
    By immersing a channel in an electrolyte solution, double layers line the interior to the channel.
    The thickness of this layer is defined by the Debye length ($l_{D}$) and is controlled the ionic concentration of the liquid.~\cite{Israelachvili2011}

    When the dimensions of a channel become small enough the double layers begin to overlap and co-ions are repelled from the channel.
    Counter-ions within a streaming cell are collected by applying a pressure gradient across the channel/cell.
    Pressure forces the liquid through the channel (from which co-ions have been removed) creating a streaming current.
    It is the streaming current that raises the voltage (streaming potential) across the cell as the ions are pumped through it.

    Streaming currents have been heavily investigated as a means of generating electrical energy from pressure gradients.~\cite{Chang2009,Daiguji2006,Daiguji2004b,Davidson2008a,Davidson2008,CherngHon2012,Jiao2014,Lu2006,Olthuis2005,Osterle1964,Pennathur2007,Ren2008a,VanderHeyden2006,Heyden2007,Xie2008,Yang2003}
    % It is expected that a channel height equal to twice the Debye length leads to maximum energy conversion efficiency.~\cite{Daiguji2004b}
    Theoretical predictions of the efficiency of standard micro/nano-fluidic channels are 2\% for pure water and 7\% for sodium chloride.~\cite{VanderHeyden2006}
    Experimental results show conversion efficiencies in the range of:
    \begin{itemize}
        \item 0.01\% by forcing water through porous glass with pore sizes from 10\thinspace--\SI{16}{\micro\metre}.~\cite{Yang2003}
        \item 0.8\% by forcing pure water through a ceramic rod populated with \SI{6}{\micro\metre} pores.~\cite{Yang2004}
        \item 3\% by forcing sodium chloride solution through a \SI{75}{\nano\metre} by \SI{50}{\micro\metre} silica channel.~\cite{Heyden2007}
        \item 0.77\% by forcing sodium chloride through a \SI{200}{\nano\metre} pore in an alumina membrane.~\cite{Lu2006}
        \item 5\% by forcing sodium chloride solution through a \SI{0.5}{\nano\metre} cylindrical pore in polyethylene terephalate foil.~\cite{Xie2008}
    \end{itemize}
    These results indicate that solutions containing salt are more efficient as a working liquid.
    And as predicted by the authors of \cite{VanderHeyden2006}, the maximum efficiency is found when the salt concentration is low.

@@

    The remainder of this work is presented as follows.
    In section \ref{sect:streamingCell} we build and measure a simple streaming potential cell.
    We confirm the relationship between pressure and applied pressure and investigate the effect of channel height.
    In section \ref{sect:waterConsumption} we estimate a typical domestic water usage profile.
    This consumption profile will be used as the basis for estimating the amount of energy available to the harvester.
    Section \ref{sect:powerRequirements} determines the amount of energy an electronic meter would need to be useful.
    Following on this, section \ref{sect:harvesterSize} presents the required dimensions a usable harvester would need.
    In section \ref{sect:results} we present the results of our measurements and quantify our predictions.
    And finally, we discuss the feasibility of the streaming cell harvester in the conclusion.

@@

    Each channel is epoxied between two acrylic reservoirs and a base plate.
    This provides a rigid structure under pressure and provides connection of wires and hoses.
    Copper wires and silicon hoses are inserted into each reservoir to facilitate voltage measurement and the application of pressure.
    A photo of one of the cells is shown in Fig. \ref{fig:cell}.

    Pressure is applied across each cell by connecting one end to a tap and leaving the other open to atmosphere.
    A Honeywell pressure sensor (model 26PC15SMT) is connected across the input/output tubes to measure the developed pressure differential.
    An Agilent precision measurement mainframe (model E5270B) is used to measure streaming potential and read the pressure sensor's output.

    \section{Estimation of Typical Household Water Consumption}:
    \label{sect:waterConsumption}
    In \cite{Heinrich2007} the author monitors the water consumption of twelve homes throughout the Kapiti Coast, New Zealand over a period of eight months.
    From that study the following metrics were found:
    \begin{itemize}
    \item Toilet
    \item Washing Machine
    \item Bath
    \item Taps
    \end{itemize}

    \section{Estimation of Power Requirements}
    \label{sect:powerRequirements}
    \section{Estimation of Streaming Cell Harvester}
    \label{sect:harvesterSize}

    \section{Results}
    \label{sect:results}

    \section{Conclusion}
    \bibliographystyle{plain}
    \bibliography{library}
